### 2021-10-04 draft.tex
Deploying to gh-pages from @ PoisotLab/ms_betadiversity_response@a7617ab9429045dcd289b2570fa69c9f37e4b2a6 🚀

\[U = 2\rho(1-p)R^2\,.\]

The part of both networks composed of overlapping species has
\(\rho pR^2\) interactions, of which \(\rho (1-q) p R^2\) are shared,
and \(\rho qp R^2\) underwent rewiring. This leads to

\[A = \rho (1-q) p R^2\,,\]

@@

\[S = \rho pq R^2\,.\]

Note that we can drop the multiplicative constant \(R^2\), making the
result independent of the size of the network. Based on these
components, we can get the values of \(\beta_{os}\) and \(\beta_{wn}\),
as presented in fig.~\ref{fig:numexp2}.

\begin{figure}
\hypertarget{fig:numexp2}{%
\centering
\includegraphics{figures/numexp2.png}
\caption{Response of \(\beta_{os}\) and \(\beta_{wn}\), and the
consequences on \(\beta_{st}\), to changes in rewiring probability
(\(q\)) and probability of species sharing (\(p\)). As expected,
\(\beta_{os}\) is not affected by species turnover, but increases with
the rewiring probability. By contrast, \(\beta_{wn}\) increases when the
rewiring probability is higher \emph{and} when fewer species are shared.
This has important consequences for \(\beta_{st}\): its value is
maximized for low species sharing, and decreases for high rewiring
probability.}\label{fig:numexp2}
}
\end{figure}

The value of \(\beta_{os}\) is entirely unchanged by variations in \(p\)
(species sharing), and responds \emph{only} to changes in \(q\) (the
probability of rewiring), whereas as expected, \(\beta_{wn}\) responded
to changes in both of these parameters: the most dissimilar networks
have low species sharing (interactions are dissimilar because brought by
unique species), and high rewiring (shared species do not share
interactions). The relative changes in \(\beta_{os}\) and \(\beta_{wn}\)
lead to predictable changes in \(\beta_{st}\): its value is maximized
when both rewiring \emph{and} species sharing are low. Increasing
rewiring decreases the impact of species turnover (because, for an equal
number of interactions, the dissimilarity of interactins in shared
species contributes more to \(\beta_{wn}\)); increasing the chance of
sharing species also does decrease \(\beta_{st}\), trivially because
there is no species turnover anymore. Note that when using the
correction of \(\beta_{st}/\beta_{wn}\), the effect of species turnover
is magnified for low probabilities of re-wiring.

In conclusion, this numerical experiment shows that the decomposition as
initially presented by Poisot \emph{et al.} (2012), \emph{i.e.} using
denominators that make sense from a network composition point of view,
succeeds at capturing the relative effect of turnover and rewiring. \#\#
Does the partition of network dissimilarity needs a new normalization?

@@

\begin{figure}
\hypertarget{fig:numexp3}{%
\centering
\includegraphics{figures/numexp3.png}
\caption{Effects of varying the connectance of the second network
(\(\rho_2\)) and the proportion of shared species (\(p\)) on the values
of the \(\beta\)-diversity components. As expected, \(\beta_{os}\) is
still independent of species turnover, and \(\beta_{wn}\) increases when
species turnover increases, or when the connectances become more
dissimilar. These figures have been generated with \(\rho_1 = 0.25\) and
\(q = 0.15\), and the results are qualitatively robust to changes in
these parameters.}\label{fig:numexp3}
}
\end{figure}

>>>>

As in the previous section, we can use these values to write

\[\beta_{os} = 1 - 2\frac{aq}{1+a}\,,\]

\[\beta_{wn} = 1 - 2\frac{ap^2q}{1+a}\,,\]

@@

\[\beta_{st} = 2aq\frac{(1-p^2)(1+a)}{a^2 + 2a + 1}\,.\]

\begin{figure}
\hypertarget{fig:connectance}{%
\centering
\includegraphics{figures/connectance/components.png}
\caption{Consequences of changing the ratio of connectances between two
equally species-rich networks on the decomposition of network
beta-diversity, assuming \(p = 0.8\). Networks with stronger differences
in connectance will tend to be more similar, because the differences in
number of links becomes extreme enough that the chances of all the links
in the sparser network being in the denser network
increases.}\label{fig:connectance}
}
\end{figure}

The values of these components are visualized in
fig.~\ref{fig:connectance}. The introduction of the connectance ratio
makes these expressions marginally more complex than in the case without
differences in connectance, but the noteworthy result remains that in
the presence of differences of connectance, the value of \(\beta_{os}\)
is still independent from species turnover. In fact, there is an
important conclusion to be drawn from this expression. The shared
species component is by definition square, meaning that from an actual
measurement of \(\beta_{os}\) between two networks for which we know the
connectance, noted \(\mathbf{b}_{os}\), we can get the probability of
rewiring by reorganizing the terms of
\(\mathbf{b}_{os} = 1 - 2aq/(1+a)\) as

\[q \approx \frac{(1-\mathbf{b}_{os})(a+1)}{2a}\,,\]

which gives the probability of rewiring as \(1-q\); note that this is an
\emph{approximation}, as it assumes that the connectances of the entire
network and the connectances of the shared components are the same.

\hypertarget{does-the-partition-of-network-dissimilarity-needs-a-new-normalization}{%
\subsection{Does the partition of network dissimilarity needs a new
normalization?}\label{does-the-partition-of-network-dissimilarity-needs-a-new-normalization}}

@@

\begin{figure}
\hypertarget{fig:commden}{%
\centering
\includegraphics{figures/common_denominator/components.png}
\caption{Effects of varying the connectance of the second network
(\(\rho_2\)) and the proportion of shared species (\(p\)) on the values
of the \(\beta\)-diversity components. As expected, \(\beta_{os}\) is
still independent of species turnover, and \(\beta_{wn}\) increases when
species turnover increases, or when the connectances become more
dissimilar. These figures have been generated with \(\rho_1 = 0.25\) and
\(q = 0.15\), and the results are qualitatively robust to changes in
these parameters.}\label{fig:commden}
}
\end{figure}
